---
title: "Home work 4"
author: "Volha Tryputsen"
date: "October 28, 2017"
output: 
  html_document:
    theme: cosmo
    highlight: haddock
    toc: true
    toc_float:
      collapse: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 6, fig.asp = .6, out.width = "90%")
options(knitr.table.format = "html") 
```

```{r libraries, message=FALSE}
library(tidyverse)
library(janitor) # clean names
library(knitr)
library(lubridate)
```


# Problem 2
```{r}
#dat_pols <- "../data/fivethirtyeight_datasets/pols-month.csv"
#dat_gdp <- "../data/fivethirtyeight_datasets/GDP.csv"

dat_pols <- "./pols-month.csv"
dat_gdp <- "./GDP.csv"
```

```{r data_import_cleaning_problem2, message=FALSE}
data_pols = 
  read_csv(dat_pols, col_types = "Dciiiciii") %>% 
  clean_names()

data_pols_clean = 
  data_pols %>%
  mutate(date = floor_date(mon, "month")) %>%
  mutate(president = as.character(ifelse(prez_dem=="1", "dem", "gop"))) %>%
  select(-c(mon, prez_dem, prez_gop)) %>%
  select(date, president, everything()) %>%
  arrange(date)

data_gdp = 
  read_csv(dat_gdp) %>% 
  clean_names()

data_gdp_clean = 
  data_gdp %>%
  mutate(gdp = replace(value, value == ".", NA)) %>%
  mutate(gdp = as.numeric(gdp)) %>%
  select(-value)

data_p2 = full_join(data_pols_clean, data_gdp_clean)
```

Dataset `data_pols_clean` contains `r nrow(data_pols_clean)` observations and `r ncol(data_pols_clean)` variables (`r colnames(data_pols_clean)`) related to the number of national politicians who are democratic or republican at a given month and year between `r data_pols_clean[1,"date"] %>% pull` and `r data_pols_clean[nrow(data_pols_clean), "date"] %>% pull`. During data cleaning step a new variable `president` was created which takes values `r unique(data_pols_clean$president)`. 
Existing indicator variables `prez_dem` and `prez_gop` were used to creat variable `president`. However, only `prez_dem` variable was used for coding for variable `president` because `prez_gop` had values of `2` which is not recorded in variable descrtiption and thus could be either a coding mistake or, perhapse, a coding for missing values.   
Variable `date` was also created by collapsing `mon` variable to the first date of the month.   

Dataset `data_gdp_clean`  contains `r nrow(data_gdp_clean)` observations and `r ncol(data_gdp_clean)` variables (`r colnames(data_gdp_clean)`) related to the value of GDP at a given month and year between `r data_gdp_clean[1,c("date")]` and `r data_gdp_clean[nrow(data_gdp_clean),"date"]`.

Datasets `data_pols_clean` and `data_gdp_clean` were merged by `date` variable. The resulting dataset `data_p2` contains `r nrow(data_p2)` observations and `r ncol(data_p2)` variables.     

