---
title: "Home work 4"
author: "Volha Tryputsen"
date: "October 28, 2017"
output: 
  html_document:
    theme: cosmo
    highlight: haddock
    toc: true
    toc_float:
      collapse: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 6, fig.asp = .6, out.width = "90%")
options(knitr.table.format = "html") 
```

```{r libraries, message=FALSE}
library(tidyverse)
library(janitor) # clean names
library(knitr)
library(lubridate)
```


# Problem 2
```{r}
#dat_pols <- "../data/fivethirtyeight_datasets/pols-month.csv"
#dat_gdp <- "../data/fivethirtyeight_datasets/GDP.csv"

dat_pols <- "./pols-month.csv"
dat_gdp <- "./GDP.csv"
```

```{r data_import_cleaning_problem2, message=FALSE}
data_pols = 
  read_csv(dat_pols, col_types = "Dciiiciii") %>% 
  clean_names()

data_pols_clean = 
  data_pols %>%
  mutate(date = floor_date(mon, "month")) %>%
  mutate(president = as.character(ifelse(prez_dem=="1", "dem", "gop"))) %>%
  select(-c(mon, prez_dem, prez_gop)) %>%
  select(date, president, everything()) %>%
  arrange(date)

data_gdp = 
  read_csv(dat_gdp) %>% 
  clean_names()

data_gdp_clean = 
  data_gdp %>%
  mutate(gdp = replace(value, value == ".", NA)) %>%
  mutate(gdp = as.numeric(gdp)) %>%
  select(-value)

data_p2 = full_join(data_pols_clean, data_gdp_clean)
```

Dataset `data_pols_clean` contains `r nrow(data_pols_clean)` observations and `r ncol(data_pols_clean)` variables (`r colnames(data_pols_clean)`) related to the number of national politicians who are democratic or republican at a given month and year between `r data_pols_clean[1,"date"] %>% pull` and `r data_pols_clean[nrow(data_pols_clean), "date"] %>% pull`. During data cleaning step a new variable `president` was created which takes values `r unique(data_pols_clean$president)`. 
Existing indicator variables `prez_dem` and `prez_gop` were used to creat variable `president`. However, only `prez_dem` variable was used for coding for variable `president` because `prez_gop` had values of `2` which is not recorded in variable descrtiption and thus could be either a coding mistake or, perhapse, a coding for missing values.   
Variable `date` was also created by collapsing `mon` variable to the first date of the month.   

Dataset `data_gdp_clean`  contains `r nrow(data_gdp_clean)` observations and `r ncol(data_gdp_clean)` variables (`r colnames(data_gdp_clean)`) related to the value of GDP at a given month and year between `r data_gdp_clean[1,c("date")]` and `r data_gdp_clean[nrow(data_gdp_clean),"date"]`.

Datasets `data_pols_clean` and `data_gdp_clean` were merged by `date` variable. The resulting dataset `data_p2` contains `r nrow(data_p2)` observations and `r ncol(data_p2)` variables.     

```{r message=FALSE}

data_p2%>%
  filter(!is.na(gdp) & !gdp == ".")%>% 
    select(date, president, gdp)%>%
    ggplot(aes(x= date, y = gdp))+geom_point(alpha = 0.5)+ geom_smooth(aes(group = president, 
            color =     president), se = FALSE, show.legend = FALSE)+ facet_grid(. ~president)+
            theme(legend.position = "bottom")+
            labs(title = "GDP Values for Each Party over Time",
                 x = "Time",
                 y = "GDP",
                 caption = "Figure 1") 

data_p2%>%
  filter(!is.na(gdp) & !gdp == ".")%>% 
    select(date, president, gdp)%>%
    ggplot(aes(x = date, y = gdp, color = president)) +
    geom_point(alpha = 0.5) +
    geom_smooth(se = FALSE) +
    theme(legend.position = "bottom")+
    labs(title = "Plot of GDP Values over Time",
         x = "Time",
         y = "GDP",
         caption = "Figure 2")
```

**Figure 1** and **Figure 2** shows GDP over time for both government parties (democratic and republican, or gop). Loess smoother used to approximate GOP trend over time for both parties reveals that when the governments with democratic presidents seize power, it shows the higher GDP. However, the GDP trend seems to change after 2010, but more data is needed to state that with certainty.  
